CQM Practical: MBQC in Quantomatic

### 2012-03-06 docs/cqm_practical/cqm_practical.tex
finished practical

In this model, steps 1 and 3 are more or less fixed, whereas step 2 is where the interesting stuff happens. in this picture, the quantum programs is represented by a circuit diagram. however this is not the only paradigm for quantum computing. another choice is \textit{measurement based quantum computing}. in this model, computation consists of these steps:

@@

\section*{The Z/X Calculus}

@@

\section*{Translating MBQC Program into Z/X graphs}

@@

\section*{Quantomatic Basics}

@@

>>>>

In this model, steps 1 and 3 are more or less fixed, whereas step 2 is where the interesting stuff happens. in this picture, the quantum programs is represented by a circuit diagram. however this is not the only paradigm for quantum computing. another choice is \textit{measurement based quantum computing} (MBQC). in this model, computation consists of these steps:

@@

\section{The Z/X Calculus}\label{sec:zx}

@@

\section{Translating MBQC Program into Z/X graphs}\label{sec:mbqc}

@@

\section{Quantomatic Basics}\label{sec:quanto}

@@

If we can close all critical pairs in a (terminating) rewrite system, it is possible to show that all normal forms are unique. That is, the system is \textit{confluent and terminating}.

\section*{Exercise 3: The CNOT Gate Using MBQC}

It is a well-known result that single-qubit unitaries and the CNOT gate are universal for quantum computing. In the Z/X calculus, we can write the CNOT gate as:
\begin{equation}\label{eq:cnot}
  \tikzfig{cnot}
\end{equation}

Recall the MBQC program from section~\ref{sec:mbqc}.

\medskip
\noindent{\color{blue} Use the translation from section~\ref{sec:mbqc} to translate this program into a graph in Quantomatic, for measurement outcomes $\{ M_2^0 \rightarrow 0, M_4^0 \rightarrow 0 \}$. Implement enough rules from section~\ref{sec:zx} to normalise the measurement pattern into (\ref{eq:cnot}).}
\medskip

Even if we do not get outcome $00$, we can still produce a CNOT gate by performing single-qubit corrections.

\medskip
\noindent{\color{blue} For each of the other three outcomes, find red and green phases to apply to qubits 1 and 3 to produce a CNOT gate. Show that all four cases reduce to (\ref{eq:cnot}) in Quantomatic.}
\medskip

Let $M_{i,A}^\alpha$ be a measurement labelled $A$ of the $i$-th qubit at angle $\alpha$. Let $Z_j^A$ be the identity for the outcome $0$ of the measurement labelled $A$ and a $Z_\pi$ gate for outcome $1$. Define $X_j^A$ similarly. These are called \textit{conditional corrections}.

\medskip
\noindent{\color{blue} Using conditional corrections, write an MBQC program that produces the CNOT gate, regardless of the measurement outcomes.}
\medskip

\section*{Bonus: Implementing Single Qubit Unitaries}

Any single qubit unitary can be written using three phase gates, $U = Z_\alpha X_\beta Z_\gamma$.

\medskip
\noindent{\color{blue} Write an MBQC program that implements an arbitrary single qubit unitary. Show that for several combinations of outcomes that the program reduces to a composition $Z_\alpha X_\beta Z_\gamma$ in Quantomatic.}
\medskip

Note that measurement angles can be chosen based on previous measurement outcomes. This is known as \textit{feed-forward} in MBQC. The existence of these two programs shows that MBQC is as powerful as the circuit model.
